---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# publicationsscraper

<!-- badges: start -->
<!-- badges: end -->

# Research and CRAN Publications Retrieval Package

## Overview

The `publicationsscraper` R package offers tools to retrieve research publications from Google Scholar and ORCID, along with CRAN package download statistics. It is tailored for academic researchers in the EBS department who wish to consolidate and analyze their research outputs, while also monitoring their contributions to CRAN, including publication updates and downloads.

The main functions included are:
- `find_cran_packages`: Searches for CRAN packages by an author's first and last name and returns relevant package information such as the number of downloads and last update date.
- `get_publications`: Retrieves publications from both Google Scholar and ORCID based on provided IDs.
- `get_all_publications`: Retrieves publications from multiple authors.
- `cran_all_pubs`: Combines CRAN package data for multiple authors.

## Installation

To install this package, you can clone the repository and load the package using:

```r
# Clone the repository
git clone <https://github.com/numbats/achievement-scraper.git>

# Install required packages
install.packages(c("pkgsearch", "tibble", "dplyr", "stringr", "scholar", "rorcid"))

# Load the package
devtools::load_all()

```

## Functions

## 1. `find_cran_packages`

This function retrieves information about CRAN packages authored by a specified individual. It returns a data frame containing the package name, number of downloads, the authors, and the last update date.

```{r}
find_cran_packages("Michael", "Lydeamore")
```

## 2. `get_publications`

This function retrieves research publications from Google Scholar and ORCID, combining the results into a single data frame. It allows for flexibility in querying either platform based on the availability of IDs.

```{r}
#Example 1: Retrieve publications from both ORCID and Google Scholar
get_publications("0000-0002-2140-5352", "vamErfkAAAAJ")
```

```{r}
#Example 2: Retrieve publications only from Google Scholar
get_publications(NA, "vamErfkAAAAJ")
```

```{r}
#Example 3: Retrieve publications only from ORCID
get_publications("0000-0002-2140-5352", NA)
```


## 3. `get_all_publications`

This function takes a data frame of authors, each with an ORCID ID and a Google Scholar ID, and returns their combined publications.

```{r}
authors_df <- tibble::tibble(
orcid_id = c("0000-0002-2140-5352", "0000-0002-1825-0097", NA, "0000-0001-5109-3700"),
scholar_id = c(NA, "vamErfkAAAAJ", "4bahYMkAAAAJ", NA)
)
get_all_publications(authors_df)
```

## 4. `cran_all_pubs

```{r}
cran_authors <- tibble::tibble(
  first_name = c("Michael", "Rob"),
  last_name = c("Lydeamore", "Hyndman")
)

cran_all_pubs(cran_authors)
```

## Dataset

`orcid_gsid.csv`

```{r}
orcid_gsid <- read_csv("data-raw/orcid_gsid.csv")
```


## Dependencies

This package relies on the following R packages:

- pkgsearch
- tibble
- dplyr
- stringr
- scholar
- rorcid

You can install these dependencies using install.packages()
```{r}
install.packages(c("pkgsearch", "tibble", "dplyr", "stringr", "scholar", "rorcid"))
```



