---
title: "Vignettes"
output: 
  rmarkdown::html_vignette:
    toc: true     
    toc_depth: 1
vignette: >
  %\VignetteIndexEntry{visualisations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE, 
  fig.width =8, 
  fig.height=5,
  fig.align = "center",
  out.width = "100%",
  dpi =300)
```

```{r loading libraries}
library(publicationsscraper)
library(ggplot2)
library(dplyr)
```


# Top Authors by Publications

```{r}
all_publications <- get_all_publications(orcid_gsid)
all_publications |>
  filter(lookup_name != "Jiti Gao") |>
  group_by(lookup_name) |>
  summarise(total_publications = n()) |>
  arrange(desc(total_publications)) |>
  slice_max(total_publications, n = 5) |>  
  ggplot(aes(x = reorder(lookup_name, total_publications), 
             y = total_publications)) + 
  geom_bar(stat = "identity", fill = "blue") +
  coord_flip() +
  theme_bw() +
  labs(title = "Top 5 Authors by Number of Publications",
       x = "Author",
       y = "Number of Publications") +
theme(plot.title = element_text(size = 16, face = "bold"),
      axis.title = element_text(size = 12))
```

## **Analysis**

The bar chart displays the top five authors by the number of publications. The most prolific author is Rob J Hyndman, with over 600 publications, reflecting his extensive contributions, particularly in statistical forecasting and time series analysis. Athanasios Pantelous comes next, with over 400 publications, showing a strong research output across various areas.Following closely, Robert Brooks has published around 350 works, reflecting significant contributions to his research fields. Maxwell King is the fourth author on the list, with just over 200 publications, indicating substantial academic output. Finally, Ralph Snyder completes the top five with around 200 publications, contributing meaningfully to the department’s research achievements.

# Publication Trend of Top Authors Over Time
```{r}
top_authors <- all_publications |>
  filter(lookup_name != "Jiti Gao") |>
  group_by(lookup_name) |>
  summarise(total_publications = n()) |>
  arrange(desc(total_publications)) |>
  slice_max(total_publications, n = 5) |>
  pull(lookup_name)

all_publications |>
  filter(lookup_name %in% top_authors) |>  
  group_by(lookup_name, publication_year) |> 
  summarise(total_publications = n()) |> 
  ggplot(aes(x = publication_year, y = total_publications, color = lookup_name)) +  
  geom_line(size = 1.2) +  
  theme_bw() +
  labs(title = "Publication Trend of Top Authors Over Time",
       x = "Year",
       y = "Total Publications",
       color = "Author") +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        legend.position = "bottom")
```

## **Analysis**

The line graph illustrates the publication trends of the top five authors from the Monash EBS Department over time. Athanasios Pantelous stands out with a sharp spike around 2010, where his publication count rises dramatically to nearly 100 in a single year. This surge highlights a particularly productive period in his academic career. After this peak, his output declines but remains steady, reflecting ongoing contributions to research. Rob J Hyndman demonstrates a more gradual increase in publications starting around 1995, peaking close to 2010. Although his output decreases afterward, he continues to maintain a relatively high level of productivity, signifying sustained involvement in statistical research and time series analysis. Robert Brooks shows a relatively steady pattern with consistent fluctuations and moderate peaks throughout the years, with his most active period yielding around 25 publications annually. On the other hand, Maxwell King and Ralph Snyder exhibit stable but lower publication rates throughout the timeline. Their contributions remain consistent, without major spikes, indicating steady research activity over the years.


# Publications by Year 
```{r}
get_all_publications(orcid_gsid) |>
  dplyr::filter(!is.na(publication_year)) |>
  dplyr::count(publication_year) |>
  ggplot(aes(x = as.numeric(publication_year), y = n)) +  
  geom_line(color= "red") +
   theme_bw() +
  labs(title = "Publications Over Time", x = "Year", y = "Number of Publications") +
theme(plot.title = element_text(size = 16, face = "bold"),
      axis.title = element_text(size = 12))
```

## **Analysis** 

The chart shows the overall trend in the number of publications over time across the Monash EBS department. The data shows a steady and gradual increase in publications up until the early 1990s, indicating relatively low research output during the initial years. However, from the mid-1990s onwards, there was a significant and consistent growth in the number of publications, with a sharp rise occurring around the mid-2000s to early 2010s, reaching a peak at over 300 publications. While there are some fluctuations post-2010, the general trend indicates that Monash EBS has seen substantial growth in its research output, possibly reflecting an increased emphasis on research, collaboration, and academic productivity in recent decades.


# Publications by Journal
```{r}
get_all_publications(orcid_gsid) |> 
  filter(!is.na(journal_name) & journal_name != "") |> 
  filter(!journal_name %in% c("arXiv preprint arXiv:", "Available at SSRN", "Monash University", "Monash Econometrics and Business Statistics Working Papers")) |> 
  group_by(journal_name) |> 
  summarise(total_publications = n()) |> 
  arrange(desc(total_publications)) |> 
  slice_max(total_publications, n = 10) |>  
  ggplot(aes(x = reorder(journal_name, total_publications), y = total_publications))  +
  geom_bar(stat = "identity", fill = "lightskyblue4") +
  coord_flip() +
   theme_bw() +
  labs(title = "Top 10 Journals by Number of Publications",
       x = "Journal",
       y = "Number of Publications") +
theme(plot.title = element_text(size = 16, face = "bold"),
      axis.title = element_text(size = 12))
```

## **Analysis**

The bar chart highlights the top 10 journals where Monash EBS staff frequently publish, with a strong focus on econometrics, forecasting, and statistical research. The "International Journal of Forecasting" leads in publication count, showcasing the department's emphasis on predictive analytics. The "Journal of Econometrics" follows, indicating a key interest in econometric studies. Other notable journals include the "Journal of Computational and Graphical Statistics" and "Journal of Applied Econometrics," reflecting contributions to statistical modeling and applied econometrics. The inclusion of journals like "Econometric Theory" and "European Journal of Operational Research" further illustrates a focus on both theoretical econometrics and operational research. Overall, the chart shows Monash EBS's substantial research output across these fields.


# Top Packages by Downloads 
```{r}
cran_all_pubs(orcid_gsid) |> 
  dplyr::arrange(desc(downloads)) |> 
  dplyr::top_n(10, downloads) |> 
  ggplot(aes(x = reorder(name, downloads), 
             y = downloads, 
             fill = ifelse(name == first(name), "highlight", "normal"))) +  
  geom_bar(stat = "identity") + 
  scale_fill_manual(values = c("highlight" = "gold", "normal" = "gold"), guide = FALSE) +  
  coord_flip() + 
   theme_bw() +
  labs(title = "Top 10 Packages by Downloads", 
       x = "Package Name", 
       y = "Downloads") +
theme(plot.title = element_text(size = 16, face = "bold"),
      axis.title = element_text(size = 12))
```

## **Analysis**

The bar chart shows the most downloaded R packages used by Monash EBS researchers, with "fracdiff" being the most popular by far. This suggests that many researchers rely on it for time series analysis, especially for fractional differencing. The next most common packages, "tsfeatures" and "fpp2," also point to a strong focus on time series forecasting and data features. Other packages like "rainbow," "hdrcde," "tsibbledata," and "fma" are also frequently used, showing they’re important for various data analysis tasks. The less downloaded packages, "Mcomp," "thief," and "demography," might be used for more specific or niche purposes. Overall, the chart highlights that time series analysis is a key area of interest for these researchers, which aligns well with advanced analytics and forecasting studies.











